---
title: "FYE_COVID19variants"
author: "Gabrielle Meza (A13747395)"
date: "7/13/2022"
output: pdf_document
---
# Initial Set up 

## Load Packages

Removed messages for loading to make Rmarkdown cleaner (message=FALSE)

Loaded RcolorBrewer for last step to changing colors

```{r message=FALSE}
library(ggplot2)
library(dplyr)
library(lubridate)
library(RColorBrewer)
```

## Import data

Import COVID 19 variant data. Use head to get sense of data rows/columns

```{r}
data <- read.csv("covid19_variants.csv")
head(data)
```

# Modify dataset 

### Lubridate and filter data

Use lubridate to specify that the date column in data set will use year-month-day format. 

Using data$colun name != (does not equal total) to get new dataset that does not include total values for each date. Can also use dyplr and >> % filtering if data set is more complex/ want to filter out specfic parts in columns or rows 


```{r}
data$date <- ymd(data$date)

clean.data <- data[data$variant_name != "Total", ]
```

# Using GGplot

## Getting started 

Use ggplot to make initial line graph of % of each variant over time, and modified labels. 

Assign initial plot to variable so later modifications are easier

```{r}
plot <-ggplot(clean.data, aes(x=date, y=percentage, colour = variant_name)) +
  geom_line() +
  labs (x= "",
        y="Percentage of Sequenced Specimens",
        colour="Variant",
        title = "COVID-19 Variants in California")

plot
```

## Plot Modifications

### Format dates 

Assigned 1 month interval for x-axis, and used dyplr to put in format Month, year. then view graph to ensure it worked. Assign to new variable to make next modifications easier. 

```{r}
p <- plot +
  scale_x_date(date_breaks = "1 month", date_labels =  "%b %Y")
p
```

### Visual edits 

Changing plot to make it look nice.

Edits made to address the following (in order):

* angle/ dates

* Remove minor axis lines (looks less cluttered)

* Removed legend background 

* Center graph title

* Resize text of axis (Bigger)

* Resize Legend title (Bigger)

```{r}
p1 <-p +
  theme(axis.text.x=element_text(angle=60, hjust=1, colour = "black"),
        panel.grid.minor = element_blank(),
        legend.key= element_blank(),
        plot.title = element_text(hjust = 0.5),
        axis.title = element_text(size = "12"),
        legend.title = element_text(size = "13"))
p1
```

### Custom colors

Finally, changing colors to be prettier using 'Paired' pallet from RcolorBrewer and assign final graph:)

```{r}
p.final <- p1 + scale_color_brewer(palette = "Paired")
p.final
```

